---
title: "ROracle HowTo"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ROracle HowTo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Disclaimer
A possible usage scenario for the package `ROracle` is described. The scenario focus on connection with the databases of Qualitas AG. Due to dependencies of the required OCI package from the underlying operating system, we are not using `ROracle` so far. Consult the other vignettes in this package for alternative solutions.


## Installation
The package `ROracle` is installed with

```{r eval=FALSE}
if (! "ROracle" %in% installed.packages()) 
  install.packages(pkgs = "ROracle")
```

Appearantly before being able to install `ROracle`, the `oracle-client-interface` (OCI) for the respective operating system must be installed. 


### Installing OCI
Depending on the operating system we are running on, we have to first download and install `OCI`. This can be done according to the description given in 

```
https://www.oracle.com/database/technologies/instant-client/macos-intel-x86-downloads.html#ic_osx_inst
```


## First Example Connections
According to `https://cran.r-project.org/web/packages/ROracle/ROracle.pdf` connections to exernal databases can be established using the following approach.


## Conclusion
Because OCI needs to be installed for every operating system that is in use, and since we use macos and linux, the possiblities of having a completely automated package that works out of the box is not given when using ROracle. Therefore, we quit looking into the usage of the package `ROracle`. 
